---
title: "R Episode 14 — Producing Reports with Quarto"
---

::: {.callout-note}
## Objectives
By the end of this episode you will be able to:

- Create a Quarto document that mixes prose, code, and output.
- Control code execution with chunk options.
- Render to HTML/PDF/Word reproducibly.

**Teaching:** 35 min · **Exercises:** 25 min
:::

## Why Quarto (not knitr/R Markdown)?

The original lesson taught **knitr** with R Markdown. We teach **Quarto** — the modern successor
from Posit. It's the *same* literate-programming idea (code + prose → a rendered document) but
multi-language, more capable, and it's the engine behind this very site. R Markdown still works;
Quarto is where new development happens.

::: {.callout-tip}
## Instructor note
Learners on this site are already reading Quarto output — point that out. The skill they learn
here is the same one that publishes ForgeLab lessons.
:::

## Anatomy of a .qmd

A Quarto document has a **YAML header**, **prose** (Markdown), and **code chunks**:

```` markdown
---
title: "Gapminder Report"
author: "Your Name"
format: html
---

## Life expectancy over time

```{r}
library(dplyr)
library(ggplot2)
gapminder <- readr::read_csv(here::here("data", "gapminder.csv"))
```

The dataset has `r nrow(gapminder)` rows. Below, life expectancy rises over time:

```{r}
#| label: fig-life
#| fig-cap: "Life expectancy by continent."
#| echo: false
ggplot(gapminder, aes(year, lifeExp, group = country)) +
  geom_line(alpha = 0.2) +
  facet_wrap(~ continent)
```
````

Note the **inline code** `` `r nrow(gapminder)` `` — numbers in your prose come straight from the
data and never go stale.

## Chunk options

Modern Quarto uses `#|` option comments inside the chunk:

```r
#| echo: false      # hide the code, show the output
#| warning: false   # suppress warnings
#| cache: true      # cache expensive chunks
#| fig-width: 8
```

## Render

```bash
quarto render report.qmd               # to the default format (html)
quarto render report.qmd --to pdf      # or pdf / docx
```

Because the report regenerates from code + data, updating the data and re-rendering updates every
number, table, and figure — no manual copy-paste, no drift.

::: {.callout-note collapse="true"}
## Challenge — A reproducible figure caption
Add an inline value to your report's prose that always reflects the data: the number of distinct
countries. What do you write?

### Solution

In the prose: `` This report covers `r length(unique(gapminder$country))` countries. `` — it
recomputes on every render.
:::

::: {.callout-important}
## Key points
- Quarto mixes prose, code chunks, and inline code into one reproducible document.
- Control chunks with `#|` options (`echo`, `warning`, `cache`, `fig-*`).
- `quarto render` outputs HTML/PDF/Word; re-rendering keeps results in sync with data.
:::

Next: [Episode 15 — Writing Good Software](r-15-writing-good-software.qmd).
